Flowchart of Current Algorithm

### 2009-05-18 docs/flowchart/flowchart.tex
moved the call to (rho h)_0 = avg(rho h)^{(1)} to step 3.  This way steps 4 and 8 exactly mirror each other.
making notation consistent with future paper 4.5

Here we outline the algorithm currently implemented in the code.  Our
starting point for this description are the series of papers describing
the development of the algorithm:
\begin{itemize}
\item {\em Low Mach Number Modeling of Type Ia
  Supernovae. I. Hydrodynamics,} A.~S.~Almgren, J.~B.~Bell, 
  C.~A.~Rendleman, \& M.~Zingale 2006, ApJ, 637, 922 (henceforth
  paper~I)
\item {\em Low Mach Number Modeling of Type Ia Supernovae. II. Energy
  Evolution,} A.~S.~Almgren, J.~B.~Bell, C.~A.~Rendleman, \& M.~Zingale
  2006, ApJ, 649, 927 (henceforth paper~II)
\item {\em Low Mach Number Modeling of Type Ia Supernovae. III. Reactions,}
A.~S.~Almgren, J.~B.~Bell, A.~Nonaka, \& M.~Zingale
  2008, ApJ, 684, 449 (henceforth paper~III)
\item {\em Low Mach Number Modeling of Type Ia Supernovae. IV. White Dwarf Convection,}
A.~S.~Almgren, J.~B.~Bell, A.~Nonaka, \& M.~Zingale,
  in preparation (henceforth paper~IV)
\end{itemize}
We also carry over some ideas from our small-scale low Mach number algorithm
for astrophysical flows:
\begin{itemize}
\item {\em Adaptive Low Mach Number Simulations of Nuclear Flames,}
J.~B.~Bell, M.~S.~Day, C.~A.~Rendleman, S.~E.~Woosley, \& M.~Zingale
2004, JCP, 195, 2, 677 (henceforth BDRWZ)
\end{itemize}

@@

plane-parallel and $\rho_0 = \rho_0 + {\rm{\bf Avg}}(\rho - \rho_0)$ in spherical,

@@

plane-parallel and $(\rho h)_0 = (\rho h)_0 + {\rm{\bf Avg}}[(\rho h) - (\rho h)_0]$ 
in spherical.  This means that {\bf React Base} no longer exists, and the intermediate 
states, $(\rho h)_0^{(1)}$ and $(\rho h)_0^{(2)}$, no longer have to exist.  

@@

Changes specific to spherical:

@@

\item For any quantity, $\phi$, we define $\overline{\phi} = ${\bf Avg}$(\phi)$
as the average over a layer at constant radius r, $\Omega_H$, as
\begin{equation}
\overline{\phi(r)}
= \frac{1}{\mathrm{A}(\Omega_H)}\int_{\Omega_H} \phi(r,{\bf x}) \; dA \enskip .
\end{equation}
In plane-parallel, this is as straightforward as averaging cell values at 
a particular height.  For spherical problems, see the description in 
paper IV, section 2.2 (or section \ref{Sec:Avg}).

@@

\item[plane-parallel:] We discretize the base state density evolution 
equation from paper III equation (29) without the $\etarho$ term,
\begin{equation}
\frac{\partial \rhozero}{\partial t} = - \nablab \cdotb \left( \rhozero w_0 \er \right)
\enskip ,
\label{eq:rho0upd_new}
\end{equation}
to compute the new base state density,
\begin{equation}
\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp} - \frac{\dt}{\dr} 
\left [ \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j+\myhalf} -  
         \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j-\myhalf} \right ] \enskip,
\end{equation}
where $j$ refers to the one-dimensional index in the radial direction.
The interface states, $\rho_0^{\outp,\nph}$, are computed using the
procedure described in paper II, Appendix A.\\

@@

%compute $\psi$.  The input arguments are different for plane-parallel and spherical.

@@

\item {\bf Advect Base Enthalpy}
is the process by which we update the base state enthalpy through $\dt$ in time.
The input arguments are different for plane-parallel and spherical.
\begin{description}
\item {\bf Advect Base Enthalpy Planar}$[(\rho h)_0^\inp,w_0^\inp,\psi^\inp] \rightarrow 
[(\rho h)_0^\outp]$:

@@

\item {\bf Correct Base}$[\rho_0^{\inp}, \rho^{\inp}] \rightarrow [\rho_0^{\outp}]$
is the process by which we adjust the base state density to account for
large-scale mixing.  It is also used to reset the base state enthalpy to be
the average of the full state enthalpy since we do not need to compute the complete
evolution of the base state enthalpy.
Unlike paper III where we used $\nabla\cdot\etarho$ as an adjustment to $\rho_0$, 
we instead simply set $\rho_0=\overline{\rho}$, which is analytically equivalent.
\begin{description}
\item[plane-parallel:] Set $\rho_0^{\outp} = {\bf Avg}(\rho^{\inp})$.
\item[spherical:] Set $\rho_0^{\outp} = \rho_0^{\inp} + {\bf Avg}(\rho^{\inp} - \rho_0^{\inp})$.
\end{description}

\item {\bf Compute} {\boldmath $w_0$} is the process by which we compute $w_0$.
The arguments are different for plane-parallel and spherical.
\begin{description}
\item {\bf Compute} {\boldmath $w_0$} {\bf Planar}$[\Sbar^{\inp},\gammabar^{\inp},p_0^{\inp},
\psi^{\inp}] \rightarrow [w_0^{\outp}]$:

@@

\item {\bf Compute Gravity}$[\rhozero^{\inp}] \rightarrow [g^{\outp}]$\\
is the process by which we compute gravity at both radial cell-centers and edges.
\begin{description}
\item For plane-parallel, gravity constant in space and time.
\item For spherical, we define the enclosed mass at edges, and then cell-centers:
\begin{equation}
m_{j-\myhalf} = \sum_{k=1}^j \frac{4}{3}\pi\left(r_{k-\myhalf}^3 - r_{k-\sfrac{3}{2}}^3\right)\rho_{0,k-1}^{\inp}; ~~~~~ m_j = m_{j-\myhalf} + \frac{4}{3}\pi \left(r_j^3 - r_{j-\myhalf}^3\right)\rho_{0,j}^{\inp} \enskip .
\end{equation}
Then, we define the gravity at both edges and cell-centers:
\begin{equation}
g_{j-\myhalf}^\outp = \frac{Gm_{j-\myhalf}}{r_{j-\myhalf}^2}; ~~~~~ 
g_j^\outp = \frac{Gm_j}{r_j^2}\enskip .
\end{equation}
\end{description}

@@

If this is the first step of the calculation ($n=0$), we set
\begin{equation}
S^{\myhalf,\star} = \frac{S^0 + S^1}{2} \enskip ,
\end{equation}
where $S^1$ is found through the iterative process that initializes the calculation.
Otherwise, following the method used in our small-scale low Mach number algorithm
(cite BDRWZ), we extrapolate to the half-time using the source
terms at the previous and current time levels
\begin{equation}
S^{\nph,\star} = S^n + \frac{\Delta t^n}{2} \frac{S^n - S^{n-1}}{\Delta t^{n-1}} \enskip .
\end{equation}
\item Compute
\begin{equation}
\overline{S^{\nph,\star}} = {\mathrm{\bf Avg}} \left(S^{\nph,\star}\right) \enskip .
\end{equation}
\item Compute $w_0^{n+\myhalf,\star}$:
\begin{description}
\item For plane-parallel,\\
{\bf Compute} {\boldmath $w_0$} {\bf Planar}$[\overline{S^{\nph,\star}},\gammabar^n,p_0^n,
\psi^{n-\myhalf}] \rightarrow [w_0^{n+\myhalf,\star}]$.

If we are using the volume discrepancy term, compute 
$\overline{p^{\inp}} = \overline{p(\rho,h,X_k)^n}$.
\item For spherical,\\
{\bf Compute} {\boldmath $w_0$} {\bf Spherical}$[\overline{S^{\nph,\star}},\gammabar^n,
\rhozero^n,p_0^n,\etarho^{n-\myhalf}] \rightarrow [w_0^{n+\myhalf,\star}]$.
\end{description}

@@

{\bf React State}$[\rho^n, (\rho h)^n, X_k^n, T^n, (\rho\Hext)^n, p_0^n] \rightarrow
[ \rho^{(1)},(\rho h)^{(1)},X_k^{(1)},T^{(1)},(\rho \omegadot_k)^{(1)},(\rho \Hnuc)^{(1)} ]$.

@@

  \item Compute the species edge states, $(\rho X_k)^{(1),\nph,\star}$,
  for the conservative update of $(\rho X_k)^{(1)}$. 
  Here we predict $\rho^{'(1)} = \rho^{(1)} - \rhozero^n$ and 
  $X_k^{(1)} = (\rho X_k)^{(1)} / \rho^{(1)}$ to time-centered edges to obtain 
  $\rho^{'(1),\nph,\star}$ and $X_k^{(1),\nph,\star}$ using a second-order
  unsplit Godunov procedure, as described in paper II,
  Appendix A, using $\Vb = \uadvone+w_0^{\nph,\star} \er$.  Convert the 
  perturbational edge state to a full state using
\begin{equation}
\rho^{(1),\nph,\star} = 
\rho^{'(1),\nph,\star} + \frac{\rho_0^n + \rho_0^{(2a),\star}}{2} \enskip .
\label{eq:rhoXedgestate}
\end{equation}
  In plane-parallel, we use 4th-order spatial averaging to compute $\rho_0$ on 
  radial faces.  In spherical, we first map $\rho_0$ to Cartesian cell-centers 
  using the mapping in section \ref{Sec:1D Cell-Centered to Cartesian Cell-Centered}, 
  and then use 2nd-order spatial averaging to compute $\rho_0$ on all faces.

@@

\begin{description}
\item For plane-parallel,
\begin{equation}
 \etarho^{\nph,\star} =  {\rm {\bf Avg}} \sum_k \left[ \left(\uadvone \cdotb \er + w_0^{\nph,\star}\right) (\rho X_k)^{(1),\nph,\star} \right] - w_0^{\nph,\star} \rho_0^{\nph,\star,{\rm pred}} \enskip ,
\end{equation}
\item For spherical, first construct 
$\eta_{\rho}^{{\rm cart},\nph,\star} = [\rho'(\Ubt\cdot\eb_r)]^{n+\myhalf,\star}$ using:
\begin{equation}
\eta_{\rho}^{{\rm cart},\nph,\star} = \left[\left(\frac{\rho^{(1)}+\rho^{(2),\star}}{2}\right)-\left(\frac{\rho_0^n+\rho_0^{(2a),\star}}{2}\right)\right] \sum_d \left(\uadvone \cdot \eb_d\right)n_d.
\end{equation}
Then, $\etarho^{\nph,\star}$ is the cell-centered average of $\eta_{\rho}^{{\rm cart},\nph,\star}$,
\begin{equation}
\etarho^{\nph,\star} = \overline{\eta_{\rho}^{{\rm cart},\nph,\star}}.
\end{equation}
This gives a cell-centered $\etarho^{\nph,\star}$.  To get $\etarho^{\nph,\star}$ at 
radial edges, average the two neighboring radial cell-centers.
\end{description}

@@

%{\bf Advect Base Pressure Planar}$[p_0^n, w_0^{\nph,\star}, \etarho^{n+\myhalf,\star}]
%\rightarrow [p_0^{n+1,\star},\psi^{\nph,\star}]$.
%\item For spherical,\\
%{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph,\star},\gammabar^{(1)},
%\overline{S^{\nph,\star}}, \rho^{(2),\star}, X_k^{(2),\star}]$\\
%$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
%\rightarrow [p_0^{n+1,\star},\psi^{n,\star},\psi^{\nph,\star}]$.
\item {\bf Enforce HSE}$[p_0^n,\rho_0^{n+1,\star},g^{n+1,\star}] \rightarrow [p_0^{n+1,\star}]$.

\item Compute $\psi$.
\begin{description}
\item For plane-parallel, 
\begin{equation}
\psi_j^{n+\myhalf,\star} = \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^{n+\myhalf,\star} 
+ \eta_{\rho,j+\myhalf}^{n+\myhalf,\star}\right) g \enskip .
\end{equation}
\item For spherical, first compute:
\begin{equation}
\gammabar^{(1)} = {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{(1)}, p_0^{n}, 
X_k^{(1)}\right) \right] \enskip ,
\end{equation}
\begin{equation}
\gammabar^{(2),\star} = {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{(2),\star}, p_0^{n+1,\star}, 
X_k^{(2),\star}\right) \right] \enskip .
\end{equation}
Then, define a base time and time-centered $\psi$:
\begin{equation}
\psi_j^{n,\star} = 
\left(\gammabar^{(1)} p_0^n\right)_j
\left \{ \Sbar_j^{n+\myhalf,\star} - 
\frac{1}{r_j^2} \left [ (r^2 w_0^{n+\myhalf,\star})_{j+\myhalf} -
          (r^2 w_0^{n+\myhalf,\star})_{j-\myhalf} \right ] \right \} \enskip ,
\end{equation}
\begin{eqnarray}
\psi_j^{n+\myhalf,\star} &=& 
\left(\frac{\gammabar^{(1)}+\gammabar^{(2),\star}}{2}\right)_j
\left(\frac{p_0^n+p_0^{n+1,\star}}{2}\right)_j \nonumber \\
&& \left \{ \Sbar_j^{n+\myhalf,\star} - 
\frac{1}{r_j^2} \left [ (r^2 w_0^{n+\myhalf,\star})_{j+\myhalf} -
          (r^2 w_0^{n+\myhalf,\star})_{j-\myhalf} \right ] \right \} \enskip .
\end{eqnarray}
\end{description}

\item {\bf Correct Base}$[(\rho h)_0^{n},(\rho h)^{(1)}] \rightarrow [(\rho h)_0^{n}]$.

\item Update $(\rho h)_0$.
\begin{description}
\item For plane-parallel,\\
{\bf Advect Base Enthalpy Planar}$[(\rho h)_0^{n}, w_0^{\nph,\star}, \psi^{\nph,\star}]
\rightarrow [(\rho h)_0^{n+1,\star}]$.
\item For spherical,\\
{\bf Advect Base Enthalpy Spherical}$[(\rho h)_0^{n}, w_0^{\nph,\star}, 
\psi^{n,\star},\psi^{\nph,\star}] \rightarrow [(\rho h)_0^{n+1,\star}]$.
\end{description}

\item Update the enthalpy.  Here we consider only the advective terms, 
neglecting the reaction terms and thermal diffusion terms, and solve a discretized version of
  \begin{equation}
  \frac{\partial (\rho h)}{\partial t}  = - \nablab \cdotb (\Ub \rho h)
+ \psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r} 
\enskip , \label{eq:rhohupdadv} 
  \end{equation}
For spherical flows, experience has shown that solving the analytically equivalent form,
\begin{equation}
  \frac{\partial (\rho h)}{\partial t}  = - \nablab \cdotb (\Ub \rho h)
+ \psi + \nabla \cdot (\Ubt p_0) - p_0 \nabla \cdot \Ubt  
\enskip , \label{eq:rhohupdadv2} 
\end{equation}
results in a better solution.

@@

\begin{equation}
(\rho h)^{(1),\nph,\star} = 
(\rho h)^{'(1),\nph,\star} + \frac{(\rho h)_0^n + (\rho h)_0^{n+1,\star}}{2}
\enskip .
\end{equation}
  In plane-parallel, we use 4th-order spatial averaging to compute $(\rho h)_0$
  on radial faces.  In spherical, we first map $h_0 \equiv (\rho h)_0/\rho_0$ 
  and $\rho_0$ to Cartesian cell-centers using the mapping in section 
  \ref{Sec:1D Cell-Centered to Cartesian Cell-Centered}, use 2nd-order spatial 
  averaging to compute $h_0$ and $\rho_0$ on all faces, and then multiply these 
  terms to get $(\rho h)_0$ on all faces.

@@

\begin{description}
\item For plane-parallel,

@@

\item For spherical,

@@

\item Compute $w_0^{n+\myhalf}$:
\begin{description}
\item For plane-parallel,\\
{\bf Compute} {\boldmath $w_0$} {\bf Planar}$[\overline{S^{\nph}},
\gammabar^{n+\myhalf},p_0^{n+\myhalf},\psi^{n+\myhalf,\star}]\rightarrow [w_0^{n+\myhalf}]$.

If we are using the volume discrepancy term, compute
\begin{equation}
\overline{p^{\inp}} = \frac{\overline{p(\rho,h,X_k)^n} + \overline{p(\rho,h,X_k)^{n+1,\star}}}{2}.
\end{equation}
\item For spherical,\\
{\bf Compute} {\boldmath $w_0$} {\bf Spherical}$[\overline{S^{\nph}},\gammabar^{n+\myhalf},
\rhozero^{n+\myhalf},p_0^{n+\myhalf},\etarho^{n+\myhalf}] \rightarrow [w_0^{n+\myhalf}]$.
\end{description}

@@

\begin{description}
\item For plane-parallel,
\begin{eqnarray}
(\rho h)^{(2)} &=& (\rho h)^{(1)} - \dt \; \left\{ \nablab \cdotb \left[ \left(\uadvtwo+w_0^{\nph} \er\right)  
(\rho h)^{(1),\nph} \right] \right\} \nonumber \\
&& + \frac{\dt}{2} \; \left(\uadvtwo \cdotb \er\right)
\left [ \left(\frac{\partial p_0}{\partial r} \right)^{n}
      + \left(\frac{\partial p_0}{\partial r} \right)^{n+1}  \right ] 
+ \dt \; \psi^{\nph} \enskip ,
\end{eqnarray}

>>>>

Here we outline the algorithm currently implemented in the code.  Our
starting point for this description are the series of papers describing
the development of the algorithm:
\begin{itemize}
\item {\em Low Mach Number Modeling of Type Ia
  Supernovae. I. Hydrodynamics,} A.~S.~Almgren, J.~B.~Bell, 
  C.~A.~Rendleman, \& M.~Zingale 2006, ApJ, 637, 922 (henceforth
  paper~I)
\item {\em Low Mach Number Modeling of Type Ia Supernovae. II. Energy
  Evolution,} A.~S.~Almgren, J.~B.~Bell, C.~A.~Rendleman, \& M.~Zingale
  2006, ApJ, 649, 927 (henceforth paper~II)\item {\em Low Mach Number Modeling of Type Ia Supernovae. III. Reactions,}
A.~S.~Almgren, J.~B.~Bell, A.~Nonaka, \& M.~Zingale
  2008, ApJ, 684, 449 (henceforth paper~III)
\item {\em Low Mach Number Modeling of Type Ia Supernovae. IV. White Dwarf Convection,}
A.~S.~Almgren, J.~B.~Bell, A.~Nonaka, \& M.~Zingale,
  in preparation (henceforth paper~IV)
\end{itemize}
We also carry over some ideas from our small-scale low Mach number algorithm
for astrophysical flows:
\begin{itemize}
\item {\em Adaptive Low Mach Number Simulations of Nuclear Flames,}
J.~B.~Bell, M.~S.~Day, C.~A.~Rendleman, S.~E.~Woosley, \& M.~Zingale
2004, JCP, 195, 2, 677 (henceforth BDRWZ)
\end{itemize}

@@

planar geometry and $\rho_0 = \rho_0 + {\rm{\bf Avg}}(\rho - \rho_0)$ in spherical geometry,

@@

planar geometry and $(\rho h)_0 = (\rho h)_0 + {\rm{\bf Avg}}[(\rho h) - (\rho h)_0]$ 
in spherical geometry.  This means that {\bf React Base} no longer exists, and the 
intermediate states, $(\rho h)_0^{(1)}$ and $(\rho h)_0^{(2)}$, no longer have to exist.  

@@

Changes specific to spherical geometry:

@@

\item For any quantity, $\phi$, we define $\overline{\phi} = ${\bf Avg}$(\phi)$
as the average over a layer at constant radius r, $\Omega_H$, as
\begin{equation}
\overline{\phi(r)}
= \frac{1}{\mathrm{A}(\Omega_H)}\int_{\Omega_H} \phi(r,{\bf x}) \; dA \enskip .
\end{equation}
\begin{description}
\item[planar:] We simply average cells at constant height.
\item[spherical:] See the description in paper IV, section 2.2 
(or section \ref{Sec:Avg} in these notes).
\end{description}

@@

\item[planar:] We discretize the base state density evolution 
equation from paper III equation (29) without the $\etarho$ term,
\begin{equation}
\frac{\partial \rhozero}{\partial t} = - \nablab \cdotb \left( \rhozero w_0 \er \right)
\enskip ,
\label{eq:rho0upd_new}
\end{equation}
to compute the new base state density,
\begin{equation}
\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp} - \frac{\dt}{\dr} 
\left [ \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j+\myhalf} -  
         \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j-\myhalf} \right ] \enskip,
\end{equation}
where $j$ refers to the one-dimensional index in the radial direction.
The interface states, $\rho_0^{\outp,\nph}$, are computed using the
procedure described in paper II, Appendix A.\\

@@

%compute $\psi$.  The input arguments are different for planar and spherical geometries.

@@

\item {\bf Advect Base Enthalpy}
is the process by which we update the base state enthalpy through $\dt$ in time.
The input arguments are different for planar and spherical geometries.
\begin{description}
\item {\bf Advect Base Enthalpy Planar}$[(\rho h)_0^\inp,w_0^\inp,\psi^\inp] \rightarrow 
[(\rho h)_0^\outp]$:

@@

\item {\bf Correct Base}$[\rho_0^{\inp}, \rho^{\inp}] \rightarrow [\rho_0^{\outp}]$
is the process by which we adjust the base state density to account for
large-scale mixing.  It is also used to reset the base state enthalpy to be
the average of the full state enthalpy since we do not need to compute the complete
evolution of the base state enthalpy.
Unlike paper III where we used $\nabla\cdot\etarho$ as an adjustment to $\rho_0$, 
we instead simply set $\rho_0=\overline{\rho}$, which is analytically equivalent.
\begin{description}
\item[planar:] Set $\rho_0^{\outp} = {\bf Avg}(\rho^{\inp})$.
\item[spherical:] Set $\rho_0^{\outp} = \rho_0^{\inp} + {\bf Avg}(\rho^{\inp} - \rho_0^{\inp})$.
\end{description}

\item {\bf Compute} {\boldmath $w_0$} is the process by which we compute $w_0$.
The arguments are different for planar and spherical geometries.
\begin{description}
\item {\bf Compute} {\boldmath $w_0$} {\bf Planar}$[\Sbar^{\inp},\gammabar^{\inp},p_0^{\inp},
\psi^{\inp}] \rightarrow [w_0^{\outp}]$:

@@

\item {\bf Compute Gravity}$[\rhozero^{\inp}] \rightarrow [g^{\outp}]$\\
is the process by which we compute gravity at both radial cell-centers and edges.
\begin{description}
\item [planar:] Gravity constant in space and time.
\item [spherical:] We define the enclosed mass at edges, and then cell-centers:
\begin{equation}
m_{j-\myhalf} = \sum_{k=1}^j \frac{4}{3}\pi\left(r_{k-\myhalf}^3 - r_{k-\sfrac{3}{2}}^3\right)\rho_{0,k-1}^{\inp}; ~~~~~ m_j = m_{j-\myhalf} + \frac{4}{3}\pi \left(r_j^3 - r_{j-\myhalf}^3\right)\rho_{0,j}^{\inp} \enskip .
\end{equation}
Then, we define the gravity at both edges and cell-centers:
\begin{equation}
g_{j-\myhalf}^\outp = \frac{Gm_{j-\myhalf}}{r_{j-\myhalf}^2}; ~~~~~ 
g_j^\outp = \frac{Gm_j}{r_j^2}\enskip .
\end{equation}
\end{description}

@@

If this is the first step of the calculation ($n=0$), we set
\begin{equation}
S^{\myhalf,\star} = \frac{S^0 + S^1}{2} \enskip ,
\end{equation}
where $S^1$ is found through the iterative process that initializes the calculation.
Otherwise, following the method used in our small-scale low Mach number algorithm
(cite BDRWZ), we extrapolate to the half-time using the source
terms at the previous and current time levels
\begin{equation}
S^{\nph,\star} = S^n + \frac{\Delta t^n}{2} \frac{S^n - S^{n-1}}{\Delta t^{n-1}} \enskip .
\end{equation}
\item Compute
\begin{equation}
\overline{S^{\nph,\star}} = {\mathrm{\bf Avg}} \left(S^{\nph,\star}\right) \enskip .
\end{equation}
\item Compute $w_0^{n+\myhalf,\star}$:
\begin{description}
\item For planar geometry,\\
{\bf Compute} {\boldmath $w_0$} {\bf Planar}$[\overline{S^{\nph,\star}},\gammabar^n,p_0^n,
\psi^{n-\myhalf}] \rightarrow [w_0^{n+\myhalf,\star}]$.

If we are using the volume discrepancy term, compute 
$\overline{p^{\inp}} = \overline{p(\rho,h,X_k)^n}$.
\item For spherical geometry,\\
{\bf Compute} {\boldmath $w_0$} {\bf Spherical}$[\overline{S^{\nph,\star}},\gammabar^n,
\rhozero^n,p_0^n,\etarho^{n-\myhalf}] \rightarrow [w_0^{n+\myhalf,\star}]$.
\end{description}

@@

\item {\bf React State}$[\rho^n, (\rho h)^n, X_k^n, T^n, (\rho\Hext)^n, p_0^n] \rightarrow
[ \rho^{(1)},(\rho h)^{(1)},X_k^{(1)},T^{(1)},(\rho \omegadot_k)^{(1)},(\rho \Hnuc)^{(1)} ]$.

\item {\bf Correct Base}$[(\rho h)_0^{n},(\rho h)^{(1)}] \rightarrow [(\rho h)_0^{n}]$.
This sets $(\rho h)_0$ to be the lateral average of $(\rho h)^{(1)}$

@@

  \item Compute the species edge states, $(\rho X_k)^{(1),\nph,\star}$,
  for the conservative update of $(\rho X_k)^{(1)}$. 
  Here we predict $\rho^{'(1)} = \rho^{(1)} - \rhozero^n$ and 
  $X_k^{(1)} = (\rho X_k)^{(1)} / \rho^{(1)}$ to time-centered edges to obtain 
  $\rho^{'(1),\nph,\star}$ and $X_k^{(1),\nph,\star}$ using a second-order
  unsplit Godunov procedure, as described in paper II,
  Appendix A, using $\Vb = \uadvone+w_0^{\nph,\star} \er$.  Convert the 
  perturbational edge state to a full state using
\begin{equation}
\rho^{(1),\nph,\star} = 
\rho^{'(1),\nph,\star} + \frac{\rho_0^n + \rho_0^{(2a),\star}}{2} \enskip .
\label{eq:rhoXedgestate}
\end{equation}
  In planar geometry, we use 4th-order spatial averaging to compute $\rho_0$ on 
  radial faces.  In spherical geometry, we first map $\rho_0$ to Cartesian cell-centers 
  using the mapping in section \ref{Sec:1D Cell-Centered to Cartesian Cell-Centered}, 
  and then use 2nd-order spatial averaging to compute $\rho_0$ on all faces.

@@

\begin{description}
\item For planar geometry,
\begin{equation}
 \etarho^{\nph,\star} =  {\rm {\bf Avg}} \sum_k \left[ \left(\uadvone \cdotb \er + w_0^{\nph,\star}\right) (\rho X_k)^{(1),\nph,\star} \right] - w_0^{\nph,\star} \rho_0^{\nph,\star,{\rm pred}} \enskip ,
\end{equation}
\item For spherical geometry, first construct 
$\eta_{\rho}^{{\rm cart},\nph,\star} = [\rho'(\Ubt\cdot\eb_r)]^{n+\myhalf,\star}$ using:
\begin{equation}
\eta_{\rho}^{{\rm cart},\nph,\star} = \left[\left(\frac{\rho^{(1)}+\rho^{(2),\star}}{2}\right)-\left(\frac{\rho_0^n+\rho_0^{(2a),\star}}{2}\right)\right] \sum_d \left(\uadvone \cdot \eb_d\right)n_d.
\end{equation}
Then, $\etarho^{\nph,\star}$ is the cell-centered average of $\eta_{\rho}^{{\rm cart},\nph,\star}$,
\begin{equation}
\etarho^{\nph,\star} = \overline{\eta_{\rho}^{{\rm cart},\nph,\star}}.
\end{equation}
This gives a cell-centered $\etarho^{\nph,\star}$.  To get $\etarho^{\nph,\star}$ at 
radial edges, average the two neighboring radial cell-centers.
\end{description}

@@

%{\bf Advect Base Pressure Planar}$[p_0^n, w_0^{\nph,\star}, \etarho^{n+\myhalf,\star}]
%\rightarrow [p_0^{n+1,\star},\psi^{\nph,\star}]$.
%\item For spherical geometry,\\
%{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph,\star},\gammabar^{(1)},
%\overline{S^{\nph,\star}}, \rho^{(2),\star}, X_k^{(2),\star}]$\\
%$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
%\rightarrow [p_0^{n+1,\star},\psi^{n,\star},\psi^{\nph,\star}]$.
\item {\bf Enforce HSE}$[p_0^n,\rho_0^{n+1,\star},g^{n+1,\star}] \rightarrow [p_0^{n+1,\star}]$.

\item Compute $\psi$.
\begin{description}
\item For planar geometry,
\begin{equation}
\psi_j^{n+\myhalf,\star} = \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^{n+\myhalf,\star} 
+ \eta_{\rho,j+\myhalf}^{n+\myhalf,\star}\right) g \enskip .
\end{equation}
\item For spherical geometry, first compute:
\begin{equation}
\gammabar^{(1)} = {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{(1)}, p_0^{n}, 
X_k^{(1)}\right) \right] \enskip ,
\end{equation}
\begin{equation}
\gammabar^{(2),\star} = {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{(2),\star}, p_0^{n+1,\star}, 
X_k^{(2),\star}\right) \right] \enskip .
\end{equation}
Then, define a base time and time-centered $\psi$:
\begin{equation}
\psi_j^{n,\star} = 
\left(\gammabar^{(1)} p_0^n\right)_j
\left \{ \Sbar_j^{n+\myhalf,\star} - 
\frac{1}{r_j^2} \left [ (r^2 w_0^{n+\myhalf,\star})_{j+\myhalf} -
          (r^2 w_0^{n+\myhalf,\star})_{j-\myhalf} \right ] \right \} \enskip ,
\end{equation}
\begin{eqnarray}
\psi_j^{n+\myhalf,\star} &=& 
\left(\frac{\gammabar^{(1)}+\gammabar^{(2),\star}}{2}\right)_j
\left(\frac{p_0^n+p_0^{n+1,\star}}{2}\right)_j \nonumber \\
&& \left \{ \Sbar_j^{n+\myhalf,\star} - 
\frac{1}{r_j^2} \left [ (r^2 w_0^{n+\myhalf,\star})_{j+\myhalf} -
          (r^2 w_0^{n+\myhalf,\star})_{j-\myhalf} \right ] \right \} \enskip .
\end{eqnarray}
\end{description}

\item Update $(\rho h)_0$.
\begin{description}
\item For planar geometry,\\
{\bf Advect Base Enthalpy Planar}$[(\rho h)_0^{n}, w_0^{\nph,\star}, \psi^{\nph,\star}]
\rightarrow [(\rho h)_0^{n+1,\star}]$.
\item For spherical geometry,\\
{\bf Advect Base Enthalpy Spherical}$[(\rho h)_0^{n}, w_0^{\nph,\star}, 
\psi^{n,\star},\psi^{\nph,\star}] \rightarrow [(\rho h)_0^{n+1,\star}]$.
\end{description}

\item Update the enthalpy.  Here we consider only the advective terms, 
neglecting the reaction terms and thermal diffusion terms, and solve a discretized version of
  \begin{equation}
  \frac{\partial (\rho h)}{\partial t}  = - \nablab \cdotb (\Ub \rho h)
+ \psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r} 
\enskip , \label{eq:rhohupdadv} 
  \end{equation}
For spherical geometry, experience has shown that solving the analytically equivalent form,
\begin{equation}
  \frac{\partial (\rho h)}{\partial t}  = - \nablab \cdotb (\Ub \rho h)
+ \psi + \nabla \cdot (\Ubt p_0) - p_0 \nabla \cdot \Ubt  
\enskip , \label{eq:rhohupdadv2} 
\end{equation}
results in a better solution.

@@

\begin{equation}
(\rho h)^{(1),\nph,\star} = 
(\rho h)^{'(1),\nph,\star} + \frac{(\rho h)_0^n + (\rho h)_0^{n+1,\star}}{2}
\enskip .
\end{equation}
  For planar geometry, we use 4th-order spatial averaging to compute $(\rho h)_0$
  on radial faces.  In spherical geometry, we first map $h_0 \equiv (\rho h)_0/\rho_0$ 
  and $\rho_0$ to Cartesian cell-centers using the mapping in section 
  \ref{Sec:1D Cell-Centered to Cartesian Cell-Centered}, use 2nd-order spatial 
  averaging to compute $h_0$ and $\rho_0$ on all faces, and then multiply these 
  terms to get $(\rho h)_0$ on all faces.

@@

\begin{description}
\item For planar geometry,

@@

\item For spherical geometry,

@@

\item Compute $w_0^{n+\myhalf}$:
\begin{description}
\item For planar geometry,\\
{\bf Compute} {\boldmath $w_0$} {\bf Planar}$[\overline{S^{\nph}},
\gammabar^{n+\myhalf},p_0^{n+\myhalf},\psi^{n+\myhalf,\star}]\rightarrow [w_0^{n+\myhalf}]$.

If we are using the volume discrepancy term, compute
\begin{equation}
\overline{p^{\inp}} = \frac{\overline{p(\rho,h,X_k)^n} + \overline{p(\rho,h,X_k)^{n+1,\star}}}{2}.
\end{equation}
\item For spherical geometry,\\
{\bf Compute} {\boldmath $w_0$} {\bf Spherical}$[\overline{S^{\nph}},\gammabar^{n+\myhalf},
\rhozero^{n+\myhalf},p_0^{n+\myhalf},\etarho^{n+\myhalf}] \rightarrow [w_0^{n+\myhalf}]$.
\end{description}

@@

\begin{description}
\item For planar geometry,
\begin{eqnarray}
(\rho h)^{(2)} &=& (\rho h)^{(1)} - \dt \; \left\{ \nablab \cdotb \left[ \left(\uadvtwo+w_0^{\nph} \er\right)  
(\rho h)^{(1),\nph} \right] \right\} \nonumber \\
&& + \frac{\dt}{2} \; \left(\uadvtwo \cdotb \er\right)
\left [ \left(\frac{\partial p_0}{\partial r} \right)^{n}
      + \left(\frac{\partial p_0}{\partial r} \right)^{n+1}  \right ] 
+ \dt \; \psi^{\nph} \enskip ,
\end{eqnarray}
